---
title: "A report"
author: "Liangzi"
date: "2025-02-19"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
#install.packages("tinytex")
#install.packages("_XQuartz")
#tinytex::install_tinytex()
knitr::opts_chunk$set(echo = TRUE)
```

# Prepare the R Environment


```{r}
# Load library/ packages
library(tidyverse)
```

```{r load-data}
# Load in lake ontario microbial community data
sample_and_taxon <-
  read_csv("sample_and_taxon.csv")
# Inspect the data
glimpse(sample_and_taxon)
View(sample_and_taxon)
```

# Lake Ontario Microbial Abundance Versus Temperature
```{r cells_vs_temp}
# temp on the x
# cell abundance on the y
# colored by env_group
# make it look nice 
ggplot(data = sample_and_taxon, by = "env_group", 
       aes(x = temperature, y = cells_per_ml, 
                   color = env_group)) +
    geom_point(aes(color = env_group)) +
  labs(x = "temperature", y = "Abundance") +
  #scale_color_manual(values = c("godenrodl"))
  geom_smooth(method = "lm", color = "black") +
  theme_classic() +
  theme(legend.position = "bottom",
        legend.title = element_blank())

```

The above plot shows that"
- There is a positive relationship between temperature and cell abundance.
- For example, deep samples are the coldest and habe the fewest cells.

```{r data=summary, include = FALSE}
# Hoow many samples do I have?
n_samples <-
  sample_and_taxon %>%
  # Only pick out sample_id column
  select(sample_id) %>%
  # Number of distinct names in the column
  n_distinct()

# Calculate the minimum and maximum temps
# min
minTemp <-
  sample_and_taxon %>% 
  summarize(round(min(temperature), digits = 1)) %>%
pull()  # only give me a number
# max
maxTemp <-
sample_and_taxon %>% 
  summarize(round(max(temperature), digits = 1)) %>%
pull()
  
```

The total number of samples is 'r n_samples'. For this set of samples, temperature ranged from a minimum of 'r minTemp' Celsius to a maximum of 'r maxTemp' Celsius
